---
title: "WIC Evaluation"
author: "Ben and Joe Brew"
date: "10/14/2015"
output:
    html_document:
      toc: true
      theme: united
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, error=FALSE}
# no scientific notation
options(scipen=999)

#packages
library(knitr)

# Create a hook for html captions
knit_hooks$set(htmlcap = function(before, options, envir) {
  if(!before) {
    paste('<p class="caption">',
          '<span style="font-size:0.6em">',
          'Above figure: ',
          options$htmlcap,
          '</span>',
          "</p>",
          sep="")
  }})

# basic knitr options 
opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.align = 'center')
options(xtable.comment = FALSE)
```

```{r, message=FALSE}
# SOURCE DATA 
source("read_in.R")
source('helpers.R')
```

```{r}
# Data explanation 
# ever_bf - Ever Infants  and Children < 24 mos. old of WIC Moms Ever Breastfed by Agency *
# ffy_2015 - participation in WIC by agency/county *
# first_tri_enroll -Percent of All WIC Prenatal Enrollees Certified for WIC during the First Trimester *
# first_tri_time1 - Percent of Prenatal Women Enrolling in WIC in the First Trimester and Certified during* Reporting Period
# healthy_weight - Percent of WIC children between 2 - 5 years of age having a Healthy Weight *
# high_risk - Nutrition Education Report for High Risk Clients having Certification End Date within Target Period *
# infants_fed -  Infants  at least 26 weeks old  Breastfed for at least 26 weeks by Agency  *
# infans_fed_june -  Ever Breastfed Infants  at least 6 Months of age and Breastfed for at least 26 weeks by * 
# infants_fed_full -  Infants  at least 26 weeks old  Fully Breastfed for at least 26 weeks by Agency *  
# infants_fed_full_june -  Ever Breastfed Infants  at least 6 months of age Fully Breastfed for at least 26 * weeks by County
# low_risk - Nutrition Education Report for Low Risk Clients having Certification End Date within Target Period
# non_hisp - Non-Hispanic Black Infants  and Children < 24 mos. old of WIC Moms Ever Breastfed by Agency *
# over_obese - overweight or obese childre >= 2 years old 
# percent_bf - Percent of Breastfed Infant Enrollees in WIC - Agency *
# percent_bf_june - All Infants up to 12 mos. of age Currently Being Breastfed by Agency/Site  (Data: FL WiSE as of 06/30/2015) *
# percent_ever - Infants and Children < 24 mos. old of WIC Moms Ever Breastfed by Agency *
# pre_entry - Percent of Prenatal Women Enrolling in WIC in the First Trimester *
```

\newpage

# Introduction

## Purpose of document

This is a prototype report for the continuous evaluation of WIC services and related outcomes at the Florida Department of Health in Alachua County.  

## Notes on data source

We do not currently have direct access to the raw data related to WIC services and outcomes.  Access has been requested.  Until granted, the only metrics available for analysis are those within this report, which are ingested manually from spreadsheets on the FDOH intranet and sharepoint sites.  

Ideally, direct database server access will be granted.  This access will permit both (a) less manual data munging, (b) "live" reporting and (c) more granular analyses of the data.

Given the current data access state, metrics are reported "as is".  This report is intentionally sparse in commentary, since our "methods" consisted primarily of visualizing the data in their available form (as opposed to generating potentially more relevant metrics ourselves).


\newpage    

# Participation 

Participation appears to have declined sharply in Alachua agency facilities last fall, and has remained relatively stable (65-70%) since then.



```{r}

# VISUALIZE OVERALL PARTICIPATION IN WIC BY AGENCY 
# Make percent 
ffy_2015_1$percent <- (ffy_2015_1$participation/ffy_2015_1$eligible)*100
ffy_2015_1$percent <- round(ffy_2015_1$percent, 2)


# arrange by percent
ffy_2015_1 <- arrange(ffy_2015_1, -percent)


# create indicator for alachua 
ffy_2015_1$County <- ifelse(ffy_2015_1$agency == '*Alachua', 'Alachua', 'Other Counties')

# group by column indicator
temp <- ffy_2015_1 %>%
  group_by(County, date) %>%
  summarise(per_mean = mean(percent, na.rm = T))

# create color vector
color <- ifelse(temp$County == 'Alachua', 'red', 'blue')
```


```{r, htmlcap= 'WIC Participation by agency', fig.height = 3.5}
par(mfrow = c(1,2))
ffy_2015_1 <- arrange(ffy_2015_1, date)
plot(ffy_2015_1$date, ffy_2015_1$percent, col = adjustcolor('black', alpha.f = 0.1), type = 'n',
     xlab = 'Date',
     ylab = 'Participation (% of eligibles)')
for (i in unique(ffy_2015_1$agency)){
  sub_data <- ffy_2015_1[ffy_2015_1$agency == i,]
  lines(sub_data$date, sub_data$percent, col = adjustcolor('blue', alpha.f = 0.3))
}
sub_data <- ffy_2015_1[ffy_2015_1$agency == '*Alachua',]  
lines(sub_data$date, sub_data$percent, col = adjustcolor('red', alpha.f = 0.6), lwd = 3)
legend('topright', lty = 1,
       col = adjustcolor(c('blue', 'red'), alpha.f = 0.4),
       legend = c('Other counties', 'Alachua'))

bp <- barplot(sub_data$percent, names.arg = sub_data$date, 
              xlab = '', ylab = 'Percent', cex.names = 0.4,
              col = adjustcolor('blue', alpha.f = 0.6),
              border = NA,
              las = 3,
              ylim = c(0, max(sub_data$percent) * 1.15))
text(x = bp[,1],
     y = sub_data$percent,
     col = adjustcolor('black', alpha.f = 0.4),
     labels = sub_data$percent,
     pos = 3,
     cex = 0.3)
par(mfrow = c(1,1))
# ggplot(sub_data, aes(x = date, y = percent)) +
#   geom_bar(stat = 'identity', color = 'red', alpha = 0.6) +
#   xlab('Date') +
#   ylab()
#     theme(axis.text.x  = element_text(size = 12, colour = 'black'),
#         axis.text.y = element_text(size = 12, colour = 'black'),
#         axis.title.y = element_text(size = 12)) + theme_hc()

```

On average, other agencies' counties have remained in the mid 70s, though participation appears to have declined state-wide at the end of the summer.



```{r, htmlcap= 'WIC Participation by agency (2)', fig.width = 5, fig.height = 3}

# Create line graph 
ggplot(data = temp, aes(date, per_mean, group = County, colour = County)) + 
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + 
  ylim(50, 90) + xlab('Date') + ylab('Percent Participated') + 
  theme(axis.text.x  = element_text(size = 12, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()
```



In regards to prenatal participation in WIC (as opposed to general participation), Alachua appears to be doing better, ranking in the top tercile for June pre-natal WIC enrollments.

```{r, htmlcap='Left: Percent of prenatal women enrolling in WIC in the first trimester. Right: Distribution of percent of prenatal women enrolling in WIC in the first trimester. Both as of June 2015.', fig.height = 3.5}
# VISUALIZE PRENATAL ENROLLMENT IN FIRST TRIMESTER BY AGENCY 
# make percent
pre_entry1$percent <- NA
for(i in 1:nrow(pre_entry1)){
   pre_entry1$percent[i] <- (pre_entry1$certified_first_tri[i]/pre_entry1$all_prenatals[i])*100
   pre_entry1$percent[i] <- round(pre_entry1$percent[i], 2)
  }

# arrange ever_bf1 descending by percent 
pre_entry1 <- arrange(pre_entry1, -percent)

# Create color vector to identify alachua 
color <- ifelse(pre_entry1$agency_name == 'Alachua', 'red', "blue")

# create object that is percent for Alachua 
per_alachua <- pre_entry1$percent[which(pre_entry1$agency == 'Alachua')]

# Bargraph 
bar <- ggplot(data = pre_entry1, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()
# historgram
hist <- ggplot(data = pre_entry1, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'blue', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Count') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-4, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)

```



Certifications in the first trimester largely follow statewide trends, but with a slight decline since the end of 2014 until now.


```{r, htmlcap= 'Percent of all WIC prenatal enrollees certified for WIC during the first trimester'}
# VISUALIZE PERCENT OF ALL WIC PRENATAL ENROLLEES USING first_tri_enroll

# create indicator for alachua 
first_tri_enroll1$County <- ifelse(first_tri_enroll1$agency == 'Alachua', 'Alachua', 'Other Counties')

# Group by indicator 
temp <- first_tri_enroll1 %>%
  group_by(County, date) %>%
  summarise(per_mean = mean(participation, na.rm = T))

# create color vector
color <- ifelse(temp$County == 'Alachua', 'red', 'blue')

# Create line graph 
ggplot(data = temp, aes(date, per_mean, group = County, colour = County)) + 
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + 
  ylim(50, 80) + xlab('Date') + ylab('Percent Participated') + 
  theme(axis.text.x  = element_text(size = 12, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()
  
```

\newpage

At the county (as opposed to agency level), Alachua mirrors statewide trends (variability is likely due to smaller size, not erratic enrollment).

```{r, htmlcap= 'Percent of prenatal women enrolling in WIC in the first trimester and certified during reporting by county', fig.width = 5, fig.height = 3}
## VIZUALIZE PERCENT OF PRENATAL WOMEN ENROLLING IN WIC IN THE FIRST TRIMESTER AND CERTIFIED DURING REPORTING PERIOD USING first_tri_time1

# create indicator for alachua 
first_tri_time1$County <- ifelse(first_tri_time1$county == 'Alachua', 'Alachua', 'Other Counties')

# Group by indicator 
temp <- first_tri_time1 %>%
#   group_by(County, time) %>%
    group_by(County, year_quarter) %>%
  summarise(per_mean = mean(value, na.rm = T))

# create color vector
color <- ifelse(temp$County == 'Alachua', 'red', 'blue')

# # create vector for quarter/year 
# quarter <- c("1st Quarter (2014)","2nd Quarter (2014)", "3rd Quarter (2014)" , "4th Quarter (2014)",
#              "1st Quarter (2015)", "2nd Quarter (2015)", "1st Quarter (2014)","2nd Quarter (2014)", 
#              "3rd Quarter (2014)" , "4th Quarter (2014)", "1st Quarter (2015)", "2nd Quarter (2015)")
# 
# # create vector of numbers
# number <- c(1:12)
# 
# # bind them together 
# temp <- cbind(temp, quarter, number)

# Make the levels of the year_quarter ordered for plotting's purpose
temp <- arrange(temp, year_quarter)
temp$year_quarter <- factor(temp$year_quarter,
                            levels = temp$year_quarter)

# make line graph. the time column is year and quarter although it is formated as year, month, day. 
# ggplot(data = temp, aes(reorder(quarter,number), per_mean, group = County, colour = County)) + 
ggplot(data = temp, aes(x = year_quarter, y = per_mean, group = County, colour = County)) +
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + 
  ylim(20, 80) + xlab('Date') + ylab('Percent Participated') + 
  theme(axis.text.x  = element_text(size = 7, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()


```



\newpage


# Breastfeeding 

Relative to other WIC agencies, Alachua's percentage of infants being breasfed is low - in the bottom quartile.  

```{r, htmlcap= 'Left: Percent of all infants up to 12 months of age currently being breastfed by agency. Right: Distribution of Percent of all infants up to 12 months of age currently being breastfed by agency. Both as of June 2015', fig.height = 3.5}
# VISUALIZE PERCENT OF ALL INFANTS UP TO 12 MONTHS OLD CURRENTLY BEING BF USING percent_bf_june 

# Make percent 
percent_bf1_june$percent <- (percent_bf1_june$currently_bf/percent_bf1_june$total_infants)*100
percent_bf1_june$percent <- round(percent_bf1_june$percent, 2)

# arrange percent_bf1_june descending by percent 
percent_bf1_june <- arrange(percent_bf1_june, -percent)

# Create color vector to identify alachua 
color <- ifelse(percent_bf1_june$agency_name == 'Alachua', 'red', "blue")

# create object that is percent for Alachua 
per_alachua <- percent_bf1_june$percent[which(percent_bf1_june$agency == 'Alachua')]

# Bargraph 
bar <- ggplot(data = percent_bf1_june, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()
# historgram
hist <- ggplot(data = percent_bf1_june, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'blue', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-4, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)


```

The share of infants ever having been breastfed is also low.  



```{r, htmlcap= 'Left: Percent of infants and children under 24 months ever breastfed by WIC moms by agency. Right: Distribution of infants and children under 24 months  ever breastfed by WIC moms by agency. Both as of June 2015.', fig.height = 3.5}
# VISUALIZE PERCENT OF INFANTS AND CHILDREN UNDER 24 MONTHS EVER BF USING ever_bf1

# Make percent
ever_bf1$percent <- (ever_bf1$ever_bf/ever_bf1$all_infants_children)*100 
ever_bf1$percent <- round(ever_bf1$percent, 2)

# arrange ever_bf1 descending by percent 
ever_bf1 <- arrange(ever_bf1, -percent)

# Create color vector to identify alachua 
color <- ifelse(ever_bf1$agency == 'Alachua', 'red', "blue")

# create object that is percent for Alachua 
per_alachua <- ever_bf1$percent[which(ever_bf1$agency == 'Alachua')]

# barplot
bar <- ggplot(data = ever_bf1, aes(reorder(agency, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()

# Histogram of Percent 
hist <- ggplot(data = ever_bf1, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'blue', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-1, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)

```


This trend remains largely unchanged over the last year.  



```{r, htmlcap= 'Percent of infants and children under 24 months ever breastfed by WIC moms by agency (2014 - 2015.', fig.height = 3, fig.width = 5}
# Through time, use percent_ever 
# First create and indicator for Alachua and group by indicator 
percent_ever1$County <- ifelse(percent_ever1$agency_name == 'ALACHUA', 'Alachua', 'Other Counties')

temp <- percent_ever1 %>%
  group_by(County, date) %>%
  summarise(per_avg = mean(percent, na.rm = T))

ggplot(data = temp, aes(date, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(60, 90) + xlab('Date') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 10, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()
  

```


As with other breastfeeding metrics, the percentage of Alachua agency infants who are breastfed per the WHO's recommendations (6 months) is much lower than the state average.   




```{r, htmlcap= 'Left: Percent of infants that were ever breastfed for at least 26 weeks and at least 6 months of age. Right: Distribution of Percent of infants that were ever breastfed for at least 26 weeks and at least 6 months of age. Both as of June 2015.', fig.height = 3.5}

# VISUALIZE PERCENT OF INFANTS THAT WERE EVER BREASTFED FOR AT LEAST 26 WEEKS AND AT LEAST 6 MONTHS OF AGE USING infants_fed1_june

# Make percent
infants_fed1_june$percent <- (infants_fed1_june$bf_infants/infants_fed1_june$all_infants)*100 
infants_fed1_june$percent <- round(infants_fed1_june$percent, 2)

# arrange infants_fed1_june descending by percent 
infants_fed1_june <- arrange(infants_fed1_june, -percent)

# Create color vector to identify alachua 
color <- ifelse(infants_fed1_june$agency_name == 'Alachua', 'red', "blue")

# create object that is percent for Alachua 
per_alachua <- infants_fed1_june$percent[which(infants_fed1_june$agency_name == 'Alachua')]

# barplot
bar <- ggplot(data = infants_fed1_june, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()

# Histogram of Percent 
hist <- ggplot(data = infants_fed1_june, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'blue', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-1, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)
```
   
Little has changed in regards to the 6-month breastfeeding rate over the last year, though Alachua appears to be making steady progress in the period since January 2015.    



```{r, htmlcap= 'Percent of infants at least 26 weeks old who have breastfed for at least 26 weeks by agency (2014-2015)', fig.height = 3, fig.width = 5}
# VISUALIZE INFANTS AT LEAST 26 WEEKS OLD WHO HAVE BF FOR AT LEAST 26 WEEKS OVER TIME using infants_fed1

# First create and indicator for Alachua and group by indicator 
infants_fed1$County <- ifelse(infants_fed1$agency_name == 'Alachua', 'Alachua', 'Other Counties')

temp <- infants_fed1 %>%
  group_by(County, date) %>%
  summarise(per_avg = mean(value, na.rm = T))

ggplot(data = temp, aes(date, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(0, 50) + xlab('Date') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 10, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()


```
 
In the distribution of 6 month breastfeeding rates, Alachua sits squarely in the middle.   




```{r, htmlcap= 'Left: Percent of infants at least 6 months of age who have fully breasfed for at least 26 weeks. Right: Distribution of percent of infants at least 6 months of age who have fully breasfed for at least 26 weeks. Both as of June 2015', fig.height = 3.5}

# Make percent
infants_full1_june$percent <- (infants_full1_june$full_bf/infants_full1_june$ever_bf)*100 
infants_full1_june$percent <- round(infants_full1_june$percent, 2)

# arrange infants_full1_june descending by percent 
infants_full1_june <- arrange(infants_full1_june, -percent)

# Create color vector to identify alachua 
color <- ifelse(infants_full1_june$agency_name == 'Alachua', 'red', "blue")

# create object that is percent for Alachua 
per_alachua <- infants_full1_june$percent[which(infants_full1_june$agency_name == 'Alachua')]

# barplot
bar <- ggplot(data = infants_full1_june, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()

# Histogram of Percent 
hist <- ggplot(data = infants_full1_june, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'blue', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-1, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)

```
  
  
  
Oddly, FDOH has multiple data sources on the 6 month breastfeeding metric (intranet vs sharepoint).  Though similar, the data are slightly different depending on the source:   





```{r, htmlcap= 'Percent of infants at least 6 months of age who have fully breasfed for at least 26 weeks (2014-2015)', fig.height = 3, fig.width = 5}

# VISUALIZE INFANTS AT LEAST 26 WEEKS OLD FULLY BREASTFED FOR AT LEAST 26 WEEKS BY AGENCY 
infants_full2$County <- ifelse(infants_full2$agency_name == 'Alachua', 'Alachua', 'Other Counties')

temp <- infants_full2 %>%
  group_by(County, date) %>%
  summarise(per_avg = mean(per_full_bf, na.rm = T))

ggplot(data = temp, aes(date, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(0, 35) + xlab('Date') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 10, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()
```


```{r, htmlcap= 'Right: Percent of Non-Hispanic Black infants and children less than 24 months old that were ever breastfed by WIC moms. Left: Distribution of percent of Non-Hispanic Black infants and children less than 24 months old that were ever breastfed by WIC moms. Both as of June 2015.', fig.height = 3.5}

# VISUALIZE PERCENT OF NON HISPANIC BLACK INFANTS AND CHILDREND UNDER 2 YEARS EVER BF

# Make percent
non_hisp1$percent <- (non_hisp1$ever_bf/non_hisp1$all)*100 
non_hisp1$percent <- round(non_hisp1$percent, 2)

# arrange non_hisp1 descending by percent 
non_hisp1 <- arrange(non_hisp1, -percent)

# Create color vector to identify alachua 
color <- ifelse(non_hisp1$agency_name == 'Alachua', 'red', "blue")

# create object that is percent for Alachua 
per_alachua <- non_hisp1$percent[which(non_hisp1$agency_name == 'Alachua')]

# barplot
bar <- ggplot(data = non_hisp1, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()

# Histogram of Percent 
hist <- ggplot(data = non_hisp1, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'blue', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-1, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)

```


\newpage

# Healthy Weight and Obesity

The percent of WIC children ages 2-5 who have a helghty weight is slightly lower in Alachua than in other counties, with a mostly flat trend.   




```{r, htmlcap= 'Percent of WIC children between ages 2-5 having a healthy weight (2014-2015)', fig.width = 5, fig.height = 3}
# VISUALIZE PERCENT OF WIC CHILDREN 2-5 WITH HEALTHY WEIGHT 
healthy_weight1$County <- ifelse(healthy_weight1$agency == 'Alachua', 'Alachua', 'Other Counties')

temp <- healthy_weight1 %>%
  group_by(County, date) %>%
  summarise(per_avg = mean(value, na.rm = T))

temp$Date <- format(temp$date, '%B\n%Y')

# arrange temp by Date
temp <- arrange(temp, date)

# relevel temp$Date with the levels being the current order of that vector
temp$Date <- factor(temp$Date, levels = temp$Date)

ggplot(data = temp, aes(Date, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(50, 80) + xlab('') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 10, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()


```

Oddly, and again showing irregularities by data source, the percent of Alachuan 2-5 WIC children who are obese appears to be significantly _lower_ than in other counties (left), while the percentage of Alachua 2-5 WIC children who are overweight also appears to be slightly _lower_.  This is an incoherent finding, given that Alachua's "healthy weight" rate is below average.    




```{r, htmlcap= 'Right: Percent of WIC children 2 years or younger that are obese by county. Left: Percent of WIC children 2 years or younger that are overweight by county.', fig.height = 3.5}
over_obese1$County <- ifelse(over_obese1$county == 'ALACHUA', 'Alachua', 'Other Counties')

temp <- over_obese1 %>%
  group_by(County, time, weight_status) %>%
  summarise(per_avg = mean(percent, na.rm = T))

# Subset by just obese and plot
obese <- temp[temp$weight_status == 'obese',]
overweight <- temp[temp$weight_status == 'overweight',]

a <- ggplot(data = obese, aes(time, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(0, 20) + xlab('') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 7, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

b <- ggplot(data = overweight, aes(time, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(0, 50) + xlab('') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 8, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

# plot side by side
multiplot(a, b, cols = 2)

```

\newpage

# Nurtrition education contacts

Frankly, we don't know what these are, so we report them as is.  Left: high-risk issuances; Right: high-risk contacts.   



```{r, htmlcap= 'Right: Average number of 2 or more high risk nurtrition education contacts. Left: Average number of 2 or more Issuances.', fig.height = 3.5}

# make county indicator 
high_risk3$County <- ifelse(high_risk3$agency_name == 'ALACHUA', 'Alachua', 'Other Counties')

# subset high_risk into 2 NE contacts and issuances 
contacts <- high_risk3[high_risk3$class == '2 or more HR NE Contacts',]
issuance <- high_risk3[high_risk3$class == '2 or more Issuances',]

# group by County for each one 
temp_contacts <- contacts %>%
  group_by(County, time) %>%
  summarise(amount = mean(value))

temp_issuances <- issuance %>%
  group_by(County, time) %>%
  summarise(amount = mean(value))

a <- ggplot(data = temp_contacts, aes(time, amount, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('')  + xlab('') + ylab('') + ylim(200, 1000) +
  theme(axis.text.x  = element_text(size = 7, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

b <- ggplot(data = temp_issuances, aes(time, amount, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + xlab('') + ylab('') + ylim(200, 1000) +
  theme(axis.text.x  = element_text(size = 8, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

# plot side by side
multiplot(a, b, cols = 2)

```


The below again shows issuances (left) and contacts (right), but this time among WIC eligibles (not just high-risk).   




```{r, htmlcap= 'Right: Percent of WIC eligible 2 or more nurtrition education contacts. Left: Percent of WIC eligible 2 or more Issuances.', fig.height = 3.5}
#Join high_risk3 and particpation from ffy_2015_1
names(ffy_2015_1)[1] <- 'agency_name'

# Recode alachua  
ffy_2015_1$agency_name <- ifelse(ffy_2015_1$agency_name == '*Alachua', 'ALACHUA', ffy_2015_1$agency_name)

# make agency_name all uppercase 
ffy_2015_1$agency_name <- toupper(ffy_2015_1$agency_name)

# keep only those with date equal to 2015-08-15 (the latest date) to avoid duplicates
temp <- ffy_2015_1[ffy_2015_1$date == '2015-08-15',]

# keep only agency name and eligible 
temp <- temp[, c("agency_name", "eligible")]

# left join by agency_name
new_high_risk <- left_join(high_risk3, temp)

# make percent 
new_high_risk$percent <- (new_high_risk$value/new_high_risk$eligible)*100 
new_high_risk$percent <- round(new_high_risk$percent, 2)

# make county indicator 
new_high_risk$County <- ifelse(new_high_risk$agency_name == 'ALACHUA', 'Alachua', 'Other Counties')

# subset high_risk into 2 NE contacts and issuances 
contacts <- new_high_risk[high_risk3$class == '2 or more HR NE Contacts',]
issuance <- new_high_risk[high_risk3$class == '2 or more Issuances',]

# group by County for each one 
temp_contacts <- contacts %>%
  group_by(County, time) %>%
  summarise(percent = mean(percent, na.rm = T))

temp_issuances <- issuance %>%
  group_by(County, time) %>%
  summarise(percent = mean(percent, na.rm = T))

a <- ggplot(data = temp_contacts, aes(time, percent, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('')  + xlab('') + ylab('Percent') +  ylim(0, 8) +
  theme(axis.text.x  = element_text(size = 7, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

b <- ggplot(data = temp_issuances, aes(time, percent, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + xlab('') + ylab('Percent') + ylim(0, 8) +
  theme(axis.text.x  = element_text(size = 8, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

# plot side by side
multiplot(a, b, cols = 2)

```

The below again shows issuances (left) and contacts (right) among the low risk population.   



```{r, htmlcap= 'Right: Average number of 2 or more low risk nurtrition education contacts. Left: Average number of 2 or more Issuances.', fig.height = 3.5}

# make county indicator 
low_risk2$County <- ifelse(low_risk2$agency_name == 'ALACHUA', 'Alachua', 'Other Counties')

# subset high_risk into 2 NE contacts and issuances 
contacts <- low_risk2[low_risk2$class == '2 or more LR NE Contacts',]
issuance <- low_risk2[low_risk2$class == '2 or more Issuances',]

# group by County for each one 
temp_contacts <- contacts %>%
  group_by(County, time) %>%
  summarise(amount = mean(value))

temp_issuances <- issuance %>%
  group_by(County, time) %>%
  summarise(amount = mean(value))

a <- ggplot(data = temp_contacts, aes(time, amount, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('')  + xlab('') + ylab('') + ylim(1000, 7000) +
  theme(axis.text.x  = element_text(size = 7, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

b <- ggplot(data = temp_issuances, aes(time, amount, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + xlab('') + ylab('') + ylim(1000, 7000) +
  theme(axis.text.x  = element_text(size = 8, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

# plot side by side
multiplot(a, b, cols = 2)

```


The below shows _multiple_ (2 or more) issuances (left) and contacts (right) among the high-risk population.   




```{r, htmlcap= 'Right: Percent of WIC eligible 2 or more high risk nurtrition education contacts. Left: Percent of WIC eligible 2 or more Issuances.', fig.height = 3.5}

# left join by agency_name
new_low_risk <- left_join(low_risk2, temp)

# make percent
new_low_risk$percent <- (new_low_risk$value/new_low_risk$eligible)*100 
new_low_risk$percent <- round(new_low_risk$percent, 2)

# make county indicator 
new_low_risk$County <- ifelse(new_low_risk$agency_name == 'ALACHUA', 'Alachua', 'Other Counties')

# subset high_risk into 2 NE contacts and issuances 
contacts <- new_low_risk[high_risk3$class == '2 or more HR NE Contacts',]
issuance <- new_low_risk[high_risk3$class == '2 or more Issuances',]

# group by County for each one 
temp_contacts <- contacts %>%
  group_by(County, time) %>%
  summarise(percent = mean(percent, na.rm = T))

temp_issuances <- issuance %>%
  group_by(County, time) %>%
  summarise(percent = mean(percent, na.rm = T))

a <- ggplot(data = temp_contacts, aes(time, percent, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('')  + xlab('') + ylab('Percent') +  ylim(0, 30) +
  theme(axis.text.x  = element_text(size = 7, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

b <- ggplot(data = temp_issuances, aes(time, percent, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + xlab('') + ylab('Percent') + ylim(0, 30) +
  theme(axis.text.x  = element_text(size = 8, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()

# plot side by side
multiplot(a, b, cols = 2)

```
