---
fig_height: 2.6
fig_width: 4
highlight: zenburn
title: "WIC Evaluation"
author: "Ben and Joe Brew"
date: "10/14/2015"
tables: yes
toc: yes
output:
  pdf_document:
    fig_caption: true
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, error=FALSE}
# no scientific notation
options(scipen=999)

#packages
library(knitr)

# basic knitr options 
opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
options(xtable.comment = FALSE)
```

```{r, message=FALSE}
# SOURCE DATA 
source("read_in.R")
source('helpers.R')
```

```{r}
# Data explanation 
# ever_bf - Ever Infants  and Children < 24 mos. old of WIC Moms Ever Breastfed by Agency *
# ffy_2015 - participation in WIC by agency/county *
# first_tri_enroll -Percent of All WIC Prenatal Enrollees Certified for WIC during the First Trimester *
# first_tri_time1 - Percent of Prenatal Women Enrolling in WIC in the First Trimester and Certified during* Reporting Period
# healthy_weight - Percent of WIC children between 2 - 5 years of age having a Healthy Weight
# high_risk - Nutrition Education Report for High Risk Clients having Certification End Date within Target Period
# infants_fed -  Infants  at least 26 weeks old  Breastfed for at least 26 weeks by Agency  *
# infans_fed_june -  Ever Breastfed Infants  at least 6 Months of age and Breastfed for at least 26 weeks by * 
# infants_fed_full -  Infants  at least 26 weeks old  Fully Breastfed for at least 26 weeks by Agency *  
# infants_fed_full_june -  Ever Breastfed Infants  at least 6 months of age Fully Breastfed for at least 26 * weeks by County
# low_risk - Nutrition Education Report for Low Risk Clients having Certification End Date within Target Period
# non_hisp - Non-Hispanic Black Infants  and Children < 24 mos. old of WIC Moms Ever Breastfed by Agency *
# over_obese - overweight or obese childre >= 2 years old 
# percent_bf - Percent of Breastfed Infant Enrollees in WIC - Agency *
# percent_bf_june - All Infants up to 12 mos. of age Currently Being Breastfed by Agency/Site  (Data: FL WiSE as of 06/30/2015) *
# percent_ever - Infants and Children < 24 mos. old of WIC Moms Ever Breastfed by Agency *
# pre_entry - Percent of Prenatal Women Enrolling in WIC in the First Trimester *
```

\newpage

# Introduction

## Purpose of document

## Notes on data source



# Participation 


```{r, fig.cap= 'WIC Participation by agency'}

# VISUALIZE OVERALL PARTICIPATION IN WIC BY AGENCY 
# Make percent 
ffy_2015_1$percent <- (ffy_2015_1$participation/ffy_2015_1$eligible)*100
ffy_2015_1$percent <- round(ffy_2015_1$percent, 2)


# arrange by percent
ffy_2015_1 <- arrange(ffy_2015_1, -percent)


# create indicator for alachua 
ffy_2015_1$County <- ifelse(ffy_2015_1$agency == '*Alachua', 'Alachua', 'Not Alachua')

# group by column indicator
temp <- ffy_2015_1 %>%
  group_by(County, date) %>%
  summarise(per_mean = mean(percent, na.rm = T))

# create color vector
color <- ifelse(temp$County == 'Alachua', 'red', 'black')

# Create line graph 
ggplot(data = temp, aes(date, per_mean, group = County, colour = County)) + 
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + 
  ylim(50, 90) + xlab('Date') + ylab('Percent Participated') + 
  theme(axis.text.x  = element_text(size = 12, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()
```




```{r, fig.cap='Left: Percent of prenatal women enrolling in WIC in the first trimester. Right: Distribution of percent of prenatal women enrolling in WIC in the first trimester. Both as of June 2015.'}
# VISUALIZE PRENATAL ENROLLMENT IN FIRST TRIMESTER BY AGENCY 
# make percent
pre_entry1$percent <- NA
for(i in 1:nrow(pre_entry1)){
   pre_entry1$percent[i] <- (pre_entry1$certified_first_tri[i]/pre_entry1$all_prenatals[i])*100
   pre_entry1$percent[i] <- round(pre_entry1$percent[i], 2)
  }

# arrange ever_bf1 descending by percent 
pre_entry1 <- arrange(pre_entry1, -percent)

# Create color vector to identify alachua 
color <- ifelse(pre_entry1$agency_name == 'Alachua', 'red', "black")

# create object that is percent for Alachua 
per_alachua <- pre_entry1$percent[which(pre_entry1$agency == 'Alachua')]

# Bargraph 
bar <- ggplot(data = pre_entry1, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()
# historgram
hist <- ggplot(data = pre_entry1, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'black', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Count') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-4, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)

```



```{r, fig.cap= 'Percent of all WIC prenatal enrollees certified for WIC during the first trimester'}
# VISUALIZE PERCENT OF ALL WIC PRENATAL ENROLLEES USING first_tri_enroll

# create indicator for alachua 
first_tri_enroll1$County <- ifelse(first_tri_enroll1$agency == 'Alachua', 'Alachua', 'Other Counties')

# Group by indicator 
temp <- first_tri_enroll1 %>%
  group_by(County, date) %>%
  summarise(per_mean = mean(participation, na.rm = T))

# create color vector
color <- ifelse(temp$County == 'Alachua', 'red', 'black')

# Create line graph 
ggplot(data = temp, aes(date, per_mean, group = County, colour = County)) + 
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + 
  ylim(50, 80) + xlab('Date') + ylab('Percent Participated') + 
  theme(axis.text.x  = element_text(size = 12, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()
  
```

```{r, fig.cap='Percent of prenatal women enrolling in WIC in the first trimester and certified during reporting by county'}
## VIZUALIZE PERCENT OF PRENATAL WOMEN ENROLLING IN WIC IN THE FIRST TRIMESTER AND CERTIFIED DURING REPORTING PERIOD USING first_tri_time1

# create indicator for alachua 
first_tri_time1$County <- ifelse(first_tri_time1$county == 'Alachua', 'Alachua', 'Other Counties')

# Group by indicator 
temp <- first_tri_time1 %>%
  group_by(County, time) %>%
  summarise(per_mean = mean(value, na.rm = T))

# create color vector
color <- ifelse(temp$County == 'Alachua', 'red', 'black')

# create vector for quarter/year 
quarter <- c("1st Quarter (2014)","2nd Quarter (2014)", "3rd Quarter (2014)" , "4th Quarter (2014)",
             "1st Quarter (2015)", "2nd Quarter (2015)", "1st Quarter (2014)","2nd Quarter (2014)", 
             "3rd Quarter (2014)" , "4th Quarter (2014)", "1st Quarter (2015)", "2nd Quarter (2015)")

# create vector of numbers
number <- c(1:12)

# bind them together 
temp <- cbind(temp, quarter, number)

# make line graph. the time column is year and quarter although it is formated as year, month, day. 
ggplot(data = temp, aes(reorder(quarter,number), per_mean, group = County, colour = County)) + 
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') + 
  ylim(20, 80) + xlab('Date') + ylab('Percent Participated') + 
  theme(axis.text.x  = element_text(size = 7, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()


```

  
```

## Breastfeeding 

```{r, fig.cap= 'Left:Percent of all infants up to 12 months of age currently being breastfed by agency (06/30/2015). Right: Distribution of Percent of all infants up to 12 months of age currently being breastfed by agency. Both as of June 2015'}
# VISUALIZE PERCENT OF ALL INFANTS UP TO 12 MONTHS OLD CURRENTLY BEING BF USING percent_bf_june 

# Make percent 
percent_bf1_june$percent <- (percent_bf1_june$currently_bf/percent_bf1_june$total_infants)*100
percent_bf1_june$percent <- round(percent_bf1_june$percent, 2)

# arrange percent_bf1_june descending by percent 
percent_bf1_june <- arrange(percent_bf1_june, -percent)

# Create color vector to identify alachua 
color <- ifelse(percent_bf1_june$agency_name == 'Alachua', 'red', "black")

# create object that is percent for Alachua 
per_alachua <- percent_bf1_june$percent[which(percent_bf1_june$agency == 'Alachua')]

# Bargraph 
bar <- ggplot(data = percent_bf1_june, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()
# historgram
hist <- ggplot(data = percent_bf1_june, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'black', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-4, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)


```

```{r, fig.cap= 'Left: Percent of infants and children under 24 months  ever breasted by WIC moms by agency. Right: Distribution of infants and children under 24 months  ever breasted by WIC moms by agency. Both as of June 2015.'}
# VISUALIZE PERCENT OF INFANTS AND CHILDREN UNDER 24 MONTHS EVER BF USING ever_bf1

# Make percent
ever_bf1$percent <- (ever_bf1$ever_bf/ever_bf1$all_infants_children)*100 
ever_bf1$percent <- round(ever_bf1$percent, 2)

# arrange ever_bf1 descending by percent 
ever_bf1 <- arrange(ever_bf1, -percent)

# Create color vector to identify alachua 
color <- ifelse(ever_bf1$agency == 'Alachua', 'red', "black")

# create object that is percent for Alachua 
per_alachua <- ever_bf1$percent[which(ever_bf1$agency == 'Alachua')]

# barplot
bar <- ggplot(data = ever_bf1, aes(reorder(agency, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()

# Histogram of Percent 
hist <- ggplot(data = ever_bf1, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'black', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-1, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)

```


```{r, fig.cap= 'Percent of infants and children under 24 months ever breasted by WIC moms by agency (2014 - 2015.'}
# Through time, use percent_ever 
# First create and indicator for Alachua and group by indicator 
percent_ever1$County <- ifelse(percent_ever1$agency_name == 'ALACHUA', 'Alachua', 'Not Alachua')

temp <- percent_ever1 %>%
  group_by(County, date) %>%
  summarise(per_avg = mean(percent, na.rm = T))

ggplot(data = temp, aes(date, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(60, 90) + xlab('Date') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 10, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()
  

```


```{r, fig.cap= 'Left: Percent of infants that were ever breastfed for at least 26 weeks and at least 6 months of age. Right: Distribution of Percent of infants that were ever breastfed for at least 26 weeks and at least 6 months of age. Both as of June 2015.'}

# VISUALIZE PERCENT OF INFANTS THAT WERE EVER BREASTFED FOR AT LEAST 26 WEEKS AND AT LEAST 6 MONTHS OF AGE USING infants_fed1_june

# Make percent
infants_fed1_june$percent <- (infants_fed1_june$bf_infants/infants_fed1_june$all_infants)*100 
infants_fed1_june$percent <- round(infants_fed1_june$percent, 2)

# arrange infants_fed1_june descending by percent 
infants_fed1_june <- arrange(infants_fed1_june, -percent)

# Create color vector to identify alachua 
color <- ifelse(infants_fed1_june$agency_name == 'Alachua', 'red', "black")

# create object that is percent for Alachua 
per_alachua <- infants_fed1_june$percent[which(infants_fed1_june$agency_name == 'Alachua')]

# barplot
bar <- ggplot(data = infants_fed1_june, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()

# Histogram of Percent 
hist <- ggplot(data = infants_fed1_june, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'black', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-1, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)
```

```{r, fig.cap= 'Percent of infants at least 26 weeks old who have breastfed for at least 26 weeks by agency (2014-2015)'}
# VISUALIZE INFANTS AT LEAST 26 WEEKS OLD WHO HAVE BF FOR AT LEAST 26 WEEKS OVER TIME using infants_fed1

# First create and indicator for Alachua and group by indicator 
infants_fed1$County <- ifelse(infants_fed1$agency_name == 'Alachua', 'Alachua', 'Not Alachua')

temp <- infants_fed1 %>%
  group_by(County, date) %>%
  summarise(per_avg = mean(value, na.rm = T))

ggplot(data = temp, aes(date, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(0, 50) + xlab('Date') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 10, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()


```
 


```{r, fig.cap= 'Left: Percent of infants at least 6 months of age who have fully breasfed for at least 26 weeks. Right: Distribution of percent of infants at least 6 months of age who have fully breasfed for at least 26 weeks. Both as of June 2015'}

# Make percent
infants_full1_june$percent <- (infants_full1_june$full_bf/infants_full1_june$ever_bf)*100 
infants_full1_june$percent <- round(infants_full1_june$percent, 2)

# arrange infants_full1_june descending by percent 
infants_full1_june <- arrange(infants_full1_june, -percent)

# Create color vector to identify alachua 
color <- ifelse(infants_full1_june$agency_name == 'Alachua', 'red', "black")

# create object that is percent for Alachua 
per_alachua <- infants_full1_june$percent[which(infants_full1_june$agency_name == 'Alachua')]

# barplot
bar <- ggplot(data = infants_full1_june, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()

# Histogram of Percent 
hist <- ggplot(data = infants_full1_june, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'black', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-1, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)

```


```{r, fig.cap= 'Percent of infants at least 6 months of age who have fully breasfed for at least 26 weeks (2014-2015)'}

# VISUALIZE INFANTS AT LEAST 26 WEEKS OLD FULLY BREASTFED FOR AT LEAST 26 WEEKS BY AGENCY 
infants_full2$County <- ifelse(infants_full2$agency_name == 'Alachua', 'Alachua', 'Not Alachua')

temp <- infants_full2 %>%
  group_by(County, date) %>%
  summarise(per_avg = mean(per_full_bf, na.rm = T))

ggplot(data = temp, aes(date, per_avg, group = County, color = County)) +  
  scale_colour_brewer(palette = 'Set1') +
  geom_line(stat = 'identity', size = 2, alpha = 0.7) + 
  ggtitle('') +
  ylim(0, 35) + xlab('Date') + ylab('Percent') + 
  theme(axis.text.x  = element_text(size = 10, colour = 'black'),
        axis.text.y = element_text(size = 12, colour = 'black'),
        axis.title.y = element_text(size = 12)) + theme_hc()
```

# non_hisp - Non-Hispanic Black Infants  and Children < 24 mos. old of WIC Moms Ever Breastfed by Agency

```{r, fig.cap= 'Right:Percent of Non-Hispanic Black infants and children less than 24 months old that were ever breastfed by WIC moms. Left: Distribution of percent of Non-Hispanic Black infants and children less than 24 months old that were ever breastfed by WIC moms. Both as of June 2015.}

# VISUALIZE PERCENT OF NON HISPANIC BLACK INFANTS AND CHILDREND UNDER 2 YEARS EVER BF

# Make percent
non_hisp1$percent <- (non_hisp1$ever_bf/non_hisp1$all)*100 
non_hisp1$percent <- round(non_hisp1$percent, 2)

# arrange non_hisp1 descending by percent 
non_hisp1 <- arrange(non_hisp1, -percent)

# Create color vector to identify alachua 
color <- ifelse(non_hisp1$agency_name == 'Alachua', 'red', "black")

# create object that is percent for Alachua 
per_alachua <- non_hisp1$percent[which(non_hisp1$agency_name == 'Alachua')]

# barplot
bar <- ggplot(data = non_hisp1, aes(reorder(agency_name, -percent), percent)) + 
  geom_bar(stat = 'identity', fill = color, alpha = 0.7, colour = 'darkgrey') + 
  xlab('') +
  ylab('Percent') + 
  ggtitle('') +
  scale_x_discrete(breaks = c("Alachua")) + 
  theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + theme_hc()

# Histogram of Percent 
hist <- ggplot(data = non_hisp1, aes(percent)) + 
  geom_histogram(binwidth = 3, fill = 'black', colour = 'grey', alpha = 0.7) + 
  geom_vline(xintercept = per_alachua, colour = 'red') + 
  ylab('Counts') + xlab('Percent') +
  ggtitle('') + 
   theme(axis.text.x  = element_text(size = 14, colour = 'black'),
        axis.text.y = element_text(size = 14, colour = 'black'),
        axis.title.y = element_text(size = 14)) + 
  geom_text(aes(per_alachua-1, 10, label = "Alachua"), colour = 'red') + theme_hc()

multiplot(bar, hist, cols = 2)

```

